Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-01 PandaEvolution.tex
Clean up and reduce 3.0/3.1

The next LHC run for taking data (Run 3) will require more resources than the
Worldwide LHC Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more
than
% 100,000
300,000 cores at over 100 Grid sites\mtnote{Addresses Alexei's comment.
Increased Grid capacity to 300K cores. Is the number of sites still 100 or
should I increase that too?}, with a peak performance of 0.3 petaFLOPS. This
capacity will be sufficient for the planned analysis and data processing, but it
will be insufficient for the Monte Carlo production workflow and any extra
activity. To alleviate these challenges, ATLAS is engaged in a program to expand
the current computing model to include additional resources such as the
opportunistic use of supercomputers as well as commercial and academic clouds.

% -----------------------------------------------------------------------------
\subsection{Use of Supercomputers with PanDA}
\label{ssec:panda-supercomputers}

@@

PanDA WMS has been designed to support distributed Grid computing. Executing
ATLAS workloads or workflows involves concurrent and/or sequential runs of
possibly large amount of jobs, each requiring no or minimal parallelization and
no runtime communication. Thus, computing infrastructure like WLCG have been
designed to aggregate large amount of computing resources across multiple sites.
While each site may deploy runtime message-passing capabilities, usually these
are not used to perform distributed computations.

There are at least two approaches to enable PanDA WMS to execute ATLAS workloads
or workflows on supercomputers: (i) using the subset of resources and
capabilities shared by both supercomputers and WLCG; (ii) reconciling the
parallel and distributed computing paradigms by means of dedicated abstractions.
The former is a pragmatic approach that enables the execution of specific
workloads by prototyping single-point solutions. The latter is a principled
approach, better suited for a production-grade solution, capable of supporting
general-purpose workloads and workflows on both supercomputers and Grid
infrastructures. These two approaches are not mutually exclusive: developing a
single-point solution gives the opportunity to better understand the problem
space, supporting the creation of abstractions for the integration of computing
paradigms.

This section illustrates the design and architecture of a job broker prototyped
by the PanDA team. This broker supports execution of part of the ATLAS
production Monte Carlo workflow on Titan, a leadership-class supercomputer
managed by the Oak Ridge Leadership Computing Facility (OLCF) at the Oak Ridge
National Laboratory (ORNL). After an analysis of the results obtained and the
lessons learned, the following section introduces the design and first
experimental characterization of a next generation executor (NGE). NGE is
designed to abstract resources and capabilities, enabling the concurrent
execution of both parallel and distributed computing on generic high performance
computing (HPC) machines.

>>>>

The next LHC run for taking data (Run 3) will require more resources than the
Worldwide LHC Computing Grid (WLCG) can provide. Currently, PanDA WMS uses more
than
% 100,000
300,000 cores at over 100 Grid sites\mtnote{Addresses Alexei's comment.
Increased Grid capacity to 300K cores. Is the number of sites still 100 or
should I increase that too?}, with a peak performance of 0.3 petaFLOPS. This
capacity will be sufficient for the planned analysis and data processing, but it
will be insufficient for the Monte Carlo production workflow and any extra
activity. To alleviate these challenges, ATLAS is engaged in a program to expand
the current computing model to include additional resources such as the
opportunistic use of supercomputers.
% as well as commercial and academic clouds.

% -----------------------------------------------------------------------------
% \subsection{Use of Supercomputers with PanDA}
% \label{ssec:panda-supercomputers}

@@

PanDA WMS has been designed to support distributed Grid computing. Executing
ATLAS workloads or workflows involves concurrent and/or sequential runs of
possibly large amount of jobs, each requiring no or minimal parallelization and
no runtime communication. Thus, computing infrastructure like WLCG have been
designed to aggregate large amount of computing resources across multiple sites.
While each site may deploy MPI capabilities, usually these are not used to
perform distributed computations.

We developed and deployed a single-point solution to better understand the
problem space and study abstractions for the integration of HTC and HPC
computing paradigms. Specifically, the PanDA team developed a job broker to
support the execution of part of the ATLAS production Monte Carlo workflow on
Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).

\mtnote{Move this to the introduction: ``There are at least two approaches to
enable PanDA WMS to execute ATLAS workloads or workflows on supercomputers: (i)
using the subset of resources and capabilities shared by both supercomputers and
WLCG; (ii) reconciling the parallel and distributed computing paradigms by means
of dedicated abstractions. The former is a pragmatic approach that enables the
execution of specific workloads by prototyping single-point solutions. The
latter is a principled approach, better suited for a production-grade solution,
capable of supporting general-purpose workloads and workflows on both
supercomputers and Grid infrastructures. \\ This section illustrates the design
and architecture of a job broker prototyped by the PanDA team. This broker
supports execution of part of the ATLAS production Monte Carlo workflow on
Titan, a leadership-class supercomputer managed by the Oak Ridge Leadership
Computing Facility (OLCF) at the Oak Ridge National Laboratory (ORNL).After an
analysis of the results obtained and the lessons learned, the following section
introduces the design and first experimental characterization of a next
generation executor (NGE). NGE is designed to abstract resources and
capabilities, enabling the concurrent execution of both parallel and distributed
computing on generic high performance computing (HPC) machines.''}
